Technology-based aids for people affected by Autism Spectrum Disorder (ASD)

### 2017-12-17 paper.tex
Revisionato report, aggiunta descrizione seconda task

Our aim is to analyze the current technology-based solutions that can help the therapist and the family of an autistic subject to interact with him.
These solutions exploit the joint use of human intelligence and artificial intelligence to improve
the powerfulness of therapies and to allow a better integration of these individuals in the society.
\end{abstract}
\begin{IEEEkeywords}
\centering
Autism Spectrum Disorder, Avatar-robots, Wearable Technology, Health Monitoring
\end{IEEEkeywords}
\clearpage

@@

% TODO Brief presentation of ASD (?)
\gls{asd} is a term used to cover a very big set of disorders. There are a lot of studies and experiments in this field, but still a lot questions about it are still not being solved \DP{Forse qui ci vanno i due punti :D}, for example, the symptoms are well known, while the causes are still mostly unknown. Nowadays \textit{screening tests} are used to classify a person as affected by \gls{asd}, but these tests have an high percentage of error (false positive or negative) and they can be administered to a patient who is at least 3 years old. For this reason, several techniques have been studied in order to establish if a children has this disorder in the first 2 years of life.

@@

\subsection{Hand-eye coordination}
\label{sec:hand_eye}
A different aspect to take into consideration when facing subjects with \gls{asd}, is the motor disorder related to the syndrome. In \cite{Casellato2017} a new approach to face the issue is presented, where the proposed protocol may be used for ad hoc subject training, helping children with ASD to enhance motor abilities related to hand-eye coordination.

@@

The actual trial consisted in two main tasks: the first task was about following a diagonal straight path moving an end-effector from a start square to an end square without exiting the path boundaries; in the second task a target moved on the screen following a curvilinear predefined path representing a digitalized version of the visuo-motor precision sub-task of the NEPSY-II protocol \DP{Cos'è il NEPSY-II protocol?}.

During the first task, a primary amount of attempts was done without applying external force disturbance to the pen-like handle; the attempts that followed, instead, were perturbed by an external force, while the final ones were not perturbed again. Results from this task show that:
\begin{itemize}
\item the task was fulfilled pretty well in the first unperturbed part;
\item during the first perturbed attempts the performance of the subjects were significantly low and attest a slow adaptation to the change;
\item after some perturbed attempts the subjects eventually manage to compensate the force almost properly, with a progressive reduction in the product $pathL \cdot T$ (where $pathL$ is the length of the path while $T$ is the time taken to travel through it);
\item when the attempts go back to be not-perturbated there is an initial overcompensation, which eventually is re-calibrated with even better performance than the ones obtained initially.
\end{itemize}

@@

During the second task there is a significant difference between \gls{td} children and children with ASD and between the first and the second attempt\DP{Il second task è strutturato come il primo? Come sono gli attempts? Non lo hai spiegato}:
\begin{itemize}
\item Even though there is not a strong correlation among the children of a same group, it can be noticed that both TD children and \gls{asd} children performed better during the second attempt, when they already knew the path to follow a priori;
\item In general, \gls{td} children had better performance than \gls{asd} children;
\item For \gls{asd} children, the \textit{hand-gaze} distance was higher than for TD children, but quite the same in both the attempts.
\end{itemize}

@@

If we unify the knowledge on the \gls{eeg} functioning principles, to the signal processing theory, we obtain a powerful diagnostic tool: representations based on Fourier Transform have been commonly applied and this brought to light that the \gls{eeg} spectrum contains five main frequency bands with characteristic waveforms and so the main way to highlight differences between normal subjects and autistic subjects can be to analyze the coherence of \gls{eeg} signals form different brain regions in order to find coupling among the signals in frequency domain.

@@

Results show that the discriminating ability of nearest neighbor classifier over STFT-BW is satisfactory (0.967 for the F3 electrode) and \autoref{tab:prediction} shows a $89.5 \%$ probability of correct classification through  STFT-BW.
Moreover, \autoref{fig:head} shows the obtained results for the three frequency bands alpha beta and gamma: coherence values that show meaningful differences with $0.01 < p < 0.05 $ are represented through dot lines, while differences with $p < 0.01$ are represented through solid lines: the resulting observation is that there are more abnormalities in connectivity for the left temporal lobe than the other ones, observation that sticks with the assumption of a correlation between \gls{asd} syndrome and abnormal behavior of temporal lobes \cite{Zilbovicius2000} \cite{ANA}.

\begin{table}
\centering
\begin{tabular}[h]{c|cc}
\toprule
\diagbox{Cases}{Membership} & Autism & Normal\\
\midrule
Autism & 90.0 & 10.0\\
Normal & 11.1 & 88.9\\
\bottomrule
\end{tabular}
\caption{Table of confusion for the correct group discrimination using STFT-BW}
\label{tab:prediction}
\end{table}

@@

To improve the performance of the sessions, the use of \gls{va} platforms has been introducted.
A \gls{va} platform consists of an Avatar, that can be a robot or a virtual character, a learning software that receives information from the Avatar, elaborates them and adjusts the response of the Avatar, and an interface for the therapist that interprets the results.
In a \gls{va} session the therapist, with the aids of the learning software, trains the Avatar to do an action that the child has to imitate. At the same time, the child tries to imitate the Avatar learning new ways to interact with the society.

@@

The first solution was proposed in \cite{Alahbabi17}, using a robotic Avatar.
They chose a robotic-based solution since it has been shown in \cite{Pioggia08} that patients manifest a more favorable response towards humanoid robots than to other non-humanoid virtual characters.

This approach has sensible advantages if compared with classical therapies, but it has also some intrinsic shortcomings:
\begin{itemize}
	\item most of the robotic or virtualized therapies are preprogrammed and this makes the session boring and repetitive;
	\item robots are limited and work in a controlled environment;
	\item usually robots do not have a quantitative behavioral analysis tool, due to the absence of therapists during their development process.
\end{itemize}

In this work they try to fulfill these lacks using a set of advanced technology tools, specifically a \textit{NAO Robot} (visible in \autoref{fig:nao_robot}) and a Perception Neuron Set of 32 sensors.

@@

During the whole process a heart-rate measuring band is worn by the partecipant, to monitor the partecipant's level of anxiety and stop or change the session whenever an intervention is required.

@@

The movements performed by the robot and imitated by the child and the therapist's questions are specifically selected to teach the child how to interact with other people and to improve his response to social stimolous.

>>>>

Our aim is to analyse the current technology-based solutions that can help the therapist and the family of an autistic subject to interact with him.
These solutions exploit the joint use of human intelligence and artificial intelligence to improve
the powerfulness of therapies and to allow a better integration of these individuals in the society.
\end{abstract}
\begin{IEEEkeywords}
\centering
Autism Spectrum Disorder, Avatar-robots, Wearable Technology, Health Monitoring
\end{IEEEkeywords}
\clearpage

@@

% TODO Brief presentation of ASD (?)
\gls{asd} is a term used to cover a very big set of disorders, there are a lot of studies and experiments in this field, but still a lot questions about it are still unsolved: for example the symptoms are well known, while the causes are still mostly unknown. Nowadays \textit{screening tests} are used to classify a person as affected by \gls{asd}, but these tests have an high percentage of error (false positive or negative) and they can be administered to a patient who is at least 3 years old. For this reason, several techniques have been studied in order to establish if a children has this disorder in the first 2 years of life.

@@

\subsection{Hand-eye coordination}
\label{sec:hand_eye}
An aspect to take into consideration when facing subjects with \gls{asd}, is the motor disorder related to the syndrome. In \cite{Casellato2017} a new approach to face the issue is presented, where the proposed protocol may be used for ad hoc subject training, helping children with ASD to enhance motor abilities related to hand-eye coordination.

@@

The actual trial consisted in two main tasks: the first task was about following a diagonal straight path moving an end-effector from a start square to an end square without exiting the path boundaries; in the second task a target moved on the screen following a curvilinear predefined path.

During the first task, a primary amount of attempts was done without applying external force disturbance to the pen-like handle; the attempts that followed, instead, were perturbed by an external force, while the final ones were not perturbed again. Results from this task show that:
\begin{itemize}
\item the task was fulfilled pretty well in the first unperturbed part;
\item during the first perturbed attempts the performance of the subjects were significantly low and attest a slow adaptation to the change;
\item after some perturbed attempts the subjects eventually manage to compensate the force almost properly, with a progressive reduction in the product $pathL \cdot T$ (where $pathL$ is the length of the path while $T$ is the time taken to travel through it);
\item when the attempts go back to be not-perturbed there is an initial overcompensation, which eventually is re-calibrated with even better performance than the ones obtained initially.
\end{itemize}

@@

The second task was composed by three trials:
\begin{itemize}
\item the participant had to catch a moving target (a grey square) with the pen-like robot handle whose feedback was represented on the screen as a red square. In this case the moving target did not leave a trace while moving;
\item this time the moving target was displayed on the screen together with its complete path, and the participant was asked to behave like in A);
\item finally the subject was asked to track the same path at a self-paced velocity, without a displayed moving target.
\end{itemize}
In all trials, the inly visual feedback that the subject was provided with, was about the end-effector position on the screen. Moreover, both gaze and robot positions were recorded.

During the second task there is a significant difference between \gls{td} children and children with ASD and between the first and the second attempt:
\begin{itemize}
\item Even though there is not a strong correlation among the children of a same group, it can be noticed that both TD children and \gls{asd} children performed better during the second attempt, when they already knew the path to follow a priori thanks to the track left behind the target;
\item In general, \gls{td} children had better performance than \gls{asd} children;
\item For \gls{asd} children, the \textit{hand-gaze} distance was higher than for TD children, but quite the same in both the attempts.
\end{itemize}

@@

If we unify the knowledge on the \gls{eeg} functioning principles to the signal processing theory, we obtain a powerful diagnostic tool. Representations based on Fourier Transform have been commonly applied and this brought to light that the \gls{eeg} spectrum contains five main frequency bands with characteristic waveforms and so the main way to highlight differences between normal subjects and autistic subjects can be to analyse the coherence of \gls{eeg} signals form different brain regions in order to find coupling among the signals in frequency domain.

@@

Results show that the discriminating ability of nearest neighbor classifier over STFT-BW is satisfactory (0.967 for the F3 electrode) and \autoref{tab:prediction} shows a $89.5 \%$ probability of correct classification through  STFT-BW.
Moreover, \autoref{fig:head} shows the obtained results for the three frequency bands alpha beta and gamma: coherence values that show meaningful differences with $0.01 < p < 0.05 $ are represented through dot lines, while differences with $p < 0.01$ are represented through solid lines: the resulting observation is that there are more abnormalities in connectivity for the left temporal lobe than the other ones. This observation sticks with the statements like the ones that Zilbovicius et al. and Alahbabi et al. made in \cite{Zilbovicius2000} \cite{ANA} about a correlation between \gls{asd} syndrome and abnormal behavior of temporal lobes.

\begin{table}
\centering
\begin{tabular}[h]{c|cc}
\toprule
\diagbox{Cases}{Membership} & Autism & Normal\\
\midrule
Autism & 89.5 & 10.0\\
Normal & 11.1 & 88.9\\
\bottomrule
\end{tabular}
\caption{Table of confusion for the correct group discrimination using STFT-BW}
\label{tab:prediction}
\end{table}

@@

To improve the performance of the sessions, the use of \gls{va} platforms has been introduced.
A \gls{va} platform consists of an Avatar, that can be a robot or a virtual character, a learning software that receives information from the Avatar, elaborates them and adjusts the response of the Avatar, and an interface for the therapist that interprets the results.
In a \gls{va} session the therapist, with the aids of the learning software, trains the Avatar to do an action that the child has to imitate. At the same time, the child tries to imitate the Avatar learning new ways to interact with the society.

@@

The first solution was proposed in \cite{Alahbabi17}, using a robotic Avatar.
They chose a robotic-based solution since it has been shown in \cite{Pioggia08} that patients manifest a more favourable response towards humanoid robots than to other non-humanoid virtual characters.

This approach has sensible advantages if compared with classical therapies, but it has also some intrinsic shortcomings:
\begin{itemize}
	\item most of the robotic or virtualized therapies are pre-programmed and this makes the session boring and repetitive;
	\item robots are limited and work in a controlled environment;
	\item usually robots do not have a quantitative behavioral analysis tool, due to the absence of therapists during their development process.
\end{itemize}

In this work they try to fulfil these lacks using a set of advanced technology tools, specifically a \textit{NAO Robot} (visible in \autoref{fig:nao_robot}) and a Perception Neuron Set of 32 sensors.

@@

During the whole process a heart-rate measuring band is worn by the participant, to monitor the participant's level of anxiety and stop or change the session whenever an intervention is required.

@@

The movements performed by the robot and imitated by the child and the therapist's questions are specifically selected to teach the child how to interact with other people and to improve his response to social stimulus.
